Enzo+Moray: Radiation Hydrodynamics Adaptive Mesh Refinement Simulations with Adaptive Ray Tracing

### 2010-12-13 mnras/mnras.tex
Massive figure filename renaming change.
Fixing spelling mistakes.

Radiation from stars and black holes strongly affects their
surroundings and plays a crucical role in topics such as stellar
atmospheres, the interstellar medium (ISM), star formation, galaxy
formation, supernovae (SNe) and cosmology.  It is a well-studied
problem \citep[e.g.][]{Mathews65, Rybicki, Mihalas84, Yorke86};
however, its treatment in multi-dimensional calculations is difficult
because of the dependence on seven variables -- three spatial, two
angular, frequency, and time.  The non-local nature of the thermal and
hydrodynamical response to radiation sources further adds to the
difficulty.  Depending on the problem of interest some simplifying
assumptions may be made.

@@

\begin{figure}
  \plotone{fig/geo_corr.eps}
  \caption{\label{fig:covering} (a) A two-dimensional illustration of
    the overlap between the beam associated with a ray $\gamma$ and a
    computational cell.  The ray has a segment length of $dr$ passing
    through the cell.  The covering area is denoted by dark grey,
    where the full area $(dr \times L_{\rm pix})$ is colored by the
    dark and light grey.  The photo-ionisation and photo-heating rates
    should be corrected by the ratio of these areas, i.e. the overlap
    fraction $f_c$.  (b) Annotation of quantities used in this
    geometric correction.}
\end{figure}

@@

In an optically thin medium, radiation is only attenuated by geometric
dilution in the local approximation to Equation (\ref{eqn:rteqn}),
i.e. the inverse square law.  With such a simple solution, the tracing
of rays is wasteful, however these rays must be propagated because the
the medium farther awap can be optically thick.  Here we describe a
method that minimizes the computational work of ray tracing in the
optically thin regime by exploiting this fact.  Each ray tracks the
total column density $N_{\rm abs}$ and the equivalent total optical
depth $\tau$ traversed by the photon.  If $\tau < \tau_{\rm thin} \sim
0.1$ after the ray exits the cell, we calculate the photo-ionisation
and photo-heating rates directly from the incoming ray instead of the
luminosity of the source.
\begin{equation}
  \label{eqn:optthin}
  k_{\rm ph} = \frac{\sigma_{\rm abs} \; P}{dt_P \; \theta_{\rm pix}}
  \; \frac{r_{\rm cell}}{r_{\rm ray}}.
\end{equation}
Note that the photon number $P$ in the ray has already been
geometrically diluted by ray splitting.  Here $r_{\rm cell}$ and
$r_{\rm ray}$ are the radii from the radiation source to the cell
centre and where the ray exits the cell.  Thus the last factor
corrects the flux to a value appropriate for the cell centre.  The
photo-heating ray is calculated in the same manner as the general
case, $\Gamma_{\rm ph} = k_{\rm ph} (E_{\rm ph} - E_i)$.  This should
only be evaluated once per cell per radiation source.  No photons are
removed from the ray.  With this method, we only require one ray
travel through each cell where the gas is optically thin, thus
reducing the computational expense.

@@

\begin{figure*}
  \epsscale{2}
  \plotone{fig/RTcoupling.eps}
  \epsscale{1}
  \caption{\label{fig:MainAlgorithm} Flow chart for the overall
    algorithm of the radiative transfer module in \enzo~that
    illustrates (1) the creation of photon packages, (2) ray tracing,
    (3) the transport of photon packages between AMR grids, and (4)
    coupling with the hydrodynamics.  The ray tracing algorithm, which
    is contained in the ``Trace Rays'' is detailed in Fig.
    \ref{fig:RTalgorithm}.}
\end{figure*}

\begin{figure*}
  \epsscale{2}
  \plotone{fig/RTalgorithm.eps}
  \epsscale{1}
  \caption{\label{fig:RTalgorithm} Flow chart for the ray tracing
    algorithm for one photon passing through a grid.  Note that only
    one step is needed in the routine to adaptively split rays.  The
    remainder is a typical ray tracing method.}
\end{figure*}

@@

\begin{figure}
  \plottwo{fig/dtp_test3noavg.eps}{fig/dtp_test3.eps}
  \caption{\label{fig:dtsmooth} Radiative transfer adaptive timestep
    in shadowing test (Test 3; \S\ref{sec:test3}) while restricting
    the neutral fraction change to 5\% in the ionisation front.  The
    unmodified timestep (top) is slightly more noisy and the minima
    are more prominent than the timestep computed with a running
    average of the last two timesteps (bottom).  The points show every
    tenth timestep taken into account for the running average.  The
    sawtooth behavior is created by the ionisation front advancing
    into the next neutral cell in the overdensity.}
\end{figure}

@@

\begin{figure}
  \plottwo{fig/test1_ionfrac.eps}{fig/test1_ifront.eps}
  \caption{\label{fig:test1_ifront} Test 1 (\hii region
    expansion with a monochromatic spectrum of 13.6 eV).  Top:
    Radially averaged profile of neutral (solid) and ionised (dashed)
    fraction at 10, 30, 100, and 500 Myr.  Bottom: Evolution of the
    calculated (top, dashed) and analytical (top, solid) Str\"{o}mgren
    radius.  The ratio of these radii are plotted in the bottom panel.}
\end{figure}

\begin{figure}
  %\epsscale{0.8}
  \plotone{fig/test1_HI500.eps}
  \caption{\label{fig:test1_HI} Test 1 (\hii region expansion
    with a monochromatic spectrum of 13.6 eV). Slice of neutral
    fraction at the origin at 500 Myr.}
  %\epsscale{1.0}
\end{figure}

@@

\begin{figure}
  \plotone{fig/test2_ionfrac.eps}
  \caption{\label{fig:test2_1a} Test 2. (\hii region expansion with a
    $T=10^5$ K blackbody spectrum).  Radially averaged profile of
    neutral (solid) and ionised (dashed) fraction at 10, 30, 100, and
    500 Myr.}
\end{figure}

\begin{figure}
  \plotone{fig/test2_neutral.eps}
  \caption{\label{fig:test2_1b} Test 2. (\hii region expansion with a
    $T=10^5$ K blackbody spectrum).  Evolution of the average neutral
    fraction.}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{fig/test2_slices}
  \epsscale{1}
  \caption{\label{fig:test2_2} Test 2. (\hii region expansion
    with a $T=10^5$ K blackbody spectrum).  Top: Slices through the
    origin of neutral fraction at 10 and 100 Myr.  Bottom: Slices of
    temperature at 10 and 100 Myr.}
\end{figure*}

\begin{figure}
  \plotone{fig/test2_ifront.eps}
  \caption{\label{fig:test2_3} Test 2. (\hii region expansion
    with a $T=10^5$ K blackbody spectrum).  Top: Evolution of the
    radius of the simulated ionisation front (dashed) and analytical
    (solid) Str\"{o}mgren radius.  Bottom: The ratio of the calculated
    and analytical Str\"{o}mgren radius.}
\end{figure}

@@

\begin{figure}
  \plottwo{fig/test3_HI.eps}{fig/test3_temp.eps}
  \caption{\label{fig:test3_1} Test 3. (I-front trapping in a dense
    clump and shadowing).  Line cut from the point source through the
    middle of the dense clump at $t = 1, 3, 5, 15$ Myr.  of the
    average neutral fraction (top) and temperature (bottom) of the
    clump.}
\end{figure}

\begin{figure}
  %\epsscale{0.85}
  \plotone{fig/test3_clump.eps}
  \caption{\label{fig:test3_2} Test 3. (I-front trapping in a dense
    clump and shadowing).  Evolution of the average ionised fraction
    (top) and temperature (bottom) of the overdense clump.}
  %\epsscale{1}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{fig/test3_slices}
  \epsscale{1}
  \caption{\label{fig:test3_3} Test 3. (I-front trapping in a dense
    clump and shadowing).  Clockwise from upper left: Slices through
    the origin of neutral fraction (1 Myr), temperature (1 Myr),
    temperature (15 Myr), and neutral fraction (15 Myr).}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.75}
  \plotone{fig/test4_ion}
  \caption{\label{fig:test4_1} Test 4. (Multiple cosmological
    sources).  Evolution of the mass- (dashed) and volume-averaged
    (solid) ionised fraction.}
  %\epsscale{1}
\end{figure}

\begin{figure*}
  %\epsscale{0.9}
  \epsscale{2}
  \plotone{fig/test4_slices}
  \epsscale{1}
  \caption{\label{fig:test4_2} Test 4. (Multiple cosmological
    sources).  Top: Slices through the origin of neutral fraction at
    50 and 200 kyr at the coordinate $z = z_{\rm box}/2$.  Bottom:
    Slices of temperature at 50 and 200 kyr.  No smoothing has been
    applied to the images.}
  %\epsscale{1}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.8}
  \plotone{fig/test5_ifront}
  \caption{\label{fig:test5_1} Test 5. (\hii region in a uniform
    medium).  Top: Growth of the computed ionisation front radius at
    an ionised fraction $x_e = 0.5$ (dashed) and at a temperature $T =
    10^4$ K (dotted) compared to the analytical estimate [solid;
    equation (\ref{eqn:test5_r})].  Bottom: The ratio of the computed
    ionisation front radii to the analytical estimate.} 
  %\epsscale{1}
\end{figure}

\begin{figure}
  \plotone{fig/test5_profiles}
  \caption{\label{fig:test5_2} Test 5. (\hii region in a uniform
    medium).  Clockwise from the upper left: Radial profiles of
    density, temperature, ionised fraction, and pressure at times $t
    =$ 10, 200, and 500 Myr.}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{fig/test5_slices}
  \epsscale{1}
  \caption{\label{fig:test5_3} Test 5. (\hii region in a uniform
    medium).  Clockwise from the upper left: Slices through the origin
    of ionised fraction, neutral fraction, temperature, and density at
    time $t =$ 500 Myr.}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.8}
  \plotone{fig/test6_ifront}
  \caption{\label{fig:test6_1} Test 6. (\hii region in a 1/$r^2$
    density profile).  Top: Growth of the computed ionisation front
    radius as $T=10^4$ K and $x_e = 0.5$ as definitions for the front.
    Bottom: Velocity of the ionisation front, computed from outputs at
    0.5 Myr intervals.  The velocity is calculated from $r_{\rm IF}$,
    whose coarse time resolution causes the noise seen in $v_{\rm
      IF}$.  It is smooth within the calculation itself.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  \plotone{fig/test6_profiles}
  \caption{\label{fig:test6_2} Test 6. (\hii region in a 1/$r^2$
    density profile).  Clockwise from the upper left: Radial profiles of
    density, temperature, ionised fraction, and pressure at times $t
    =$ 3, 10, and 25 Myr.}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{fig/test6_slices}
  \epsscale{1}
  \caption{\label{fig:test6_3} Test 6. (\hii region in a 1/$r^2$
    density profile).  Clockwise from the upper left: Slices through the origin
    of ionised fraction, neutral fraction, temperature, and density at
    time $t =$ 25 Myr.}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.9}
  \plotone{fig/test7_lines}
  \caption{\label{fig:test7_1} Test 7. (Photo-evaporation of a dense
    clump).  Line cuts from the point source through the middle of the
    dense clump at $t = 1, 10, 50$ Myr of (clockwise from the upper
    left) density, temperature, pressure, and neutral fraction.}
  %\epsscale{1}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{fig/test7_slices10}
  \epsscale{1}
  \caption{\label{fig:test7_2} Test 7. (Photo-evaporation of a dense
    clump).  Clockwise from the upper left: Slices through the clump
    centre of neutral fraction, pressure, temperature, and density at
    time $t =$ 10 Myr.}
\end{figure*}

\begin{figure*}
  \epsscale{2}
  \plotone{fig/test7_slices50}
  \epsscale{1}
  \caption{\label{fig:test7_3} Test 7. (Photo-evaporation of a dense
    clump).  Same as Fig. \ref{fig:test7_1} but at $t = 50$ Myr.}
\end{figure*}

@@

\begin{figure*}
  \epsscale{2}
  \plotone{fig/test8_slices}
  \epsscale{1}
  \caption{\label{fig:test8_1} Test 8. (Champagne flow from a dense
    clump).  Slices of density through the initial clump centre in the
    x-y plane (top) and x-z plane (bottom) at $t = 10, 40, 100, 150$
    kyr.  Notice the instabilities that grow from perturbations
    created while the \hii region is contained in the dense
    clump.}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.75}
  \plotone{fig/test9thick_rho}
  \plotone{fig/test9thick_temp}
  \plotone{fig/test9thick_efrac}
  \caption{\label{fig:test9_thick} Test 9. (Irradiated Rayleigh-Taylor
    instability; optically thick case).  Slices at $y=0$ of density
    (top), temperature (middle), and electron fraction (bottom).  The
    source turns on at $t=0$.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  %\epsscale{0.75}
  \plotone{fig/test9thin_rho}
  \plotone{fig/test9thin_temp}
  \plotone{fig/test9thin_efrac}
  \caption{\label{fig:test9_thin} Same as Fig. \ref{fig:test9_thick}
    but for the optically thin case.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure*}
  \epsscale{2}
  \plotone{fig/test10_slices}
  \epsscale{1}
  \caption{\label{fig:test10} Test 10 (Photo-evaporation of a
    blastwave).  Slices of density (top) and temperature (bottom) at
    $t = 2.5, 3, 5, 7.5$ Myr in the $x-z$ plane.  As the R-type
    ionisation front propagates through the blastwave centre,
    instabilities grow from the slightly inhomogeneous hot and rarefied
    medium.  Note that the dense shell of the blastwave also creates
    dense inward fingers in the ionisation front shock.}
  %\epsscale{1}
\end{figure*}

@@

\begin{figure*}
  %\epsscale{1}
  \epsscale{2}
  \plotone{fig/test11_slices}
  \epsscale{1}
  \caption{\label{fig:test11} Test 11 (Collimated radiation from a
    dense clump).  Slices of density (top) and temperature (bottom) at
    $t = 0.1, 3.25, 10.75, 23.25$ Myr.  The conical \hii region
    drives shocks transversely into the overdense sphere and creates
    polar champagne flows.  The ambient medium is heated to $T \sim 3
    \times 10^4$ K as the ionisation front passes the
    constant-pressure cloud surface.  The ionisation front changes
    from D-type to R-type after it enters the ambient medium.}
  %\epsscale{1}
\end{figure*}

@@

\begin{figure}
  %\epsscale{1.15}
  \plottwo{fig/test12_kph}{fig/test12_radial}
  \caption{\label{fig:test12_kph} Test 12 (Time variations of the
    source luminosity).  Top: Slice of the photo-ionisation rate
    \kph~through the origin.  The source has a duty cycle of 0.5 Myr,
    and the box has a light crossing time of 3.3 Myr.  The shells of
    high \kph~originate from radiation that was emitted when the
    source was at its peak luminosity, illustrating the
    time-dependence of the radiative transfer equation.  Bottom:
    Radial profile \kph~with the inverse square law overplotted.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  %\epsscale{0.8}
  \plotone{fig/test12_profiles}
  \caption{\label{fig:test12_profiles} Test 12 (Time variations of the
    source luminosity).  Radial profiles of (clockwise from upper
    left) density, temperature, pressure, and ionised fraction at $t =
    50, 100, 150, 200, 250$ Myr.  In this problem, the time variations
    in the source have little effect on the overall \hii region
    expansion.  Inside the ionisation front at $t = 50$ Myr, there are
    small density perturbations that are created by the variable
    source that are later smoothed out over a sound crossing time.}
  %\epsscale{1}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{fig/test14_slices_xy}
  \epsscale{1}
  \caption{\label{fig:test14_1} Test 14. (\hii region with MHD).
    Left to right: slices of density at $t = 0.18, 0.53, 1.58$ Myr in
    the x-y plane.  The streamlines show the magnetic field.}
\end{figure*}

\begin{figure*}
  \epsscale{2}
  \plotone{fig/test14_slices_yz}
  \epsscale{1}
  \caption{\label{fig:test14_2} Test 14. (\hii region with
    MHD). Slices of density (top) and the x-component of the magnetic
    field (bottom) in the y-z plane at $t = 0.18, 0.53, 1.58$ Myr
    (left to right).}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.75}
  \plotone{fig/ifront_res}
  \caption{\label{fig:dx_dep1} Growth of the ionisation front radius,
    compared to the analytical radius, in Test 1 with varying spatial
    resolutions.  At resolutions of $16^3$ and $32^3$, the ionisation
    front is underestimated for the first $\sim25$ Myr but converges
    within 0.5\% of the higher resolution runs.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure}
  %\epsscale{1}
  %\epsscale{2}
  \plotone{fig/phic_slices}
  \epsscale{1}
  \caption{\label{fig:ang_dep1} Variations in the photo-ionisation
    rates for different ray-to-cell samplings $\Phi_c$.  The colormap
    only spans a factor of 3 to enhance the contrast.  In comparison,
    the photo-ionisation rate actually spans 4 orders of magnitude in
    this test.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  %\epsscale{0.75}
  \plotone{fig/phic_dep}
  \caption{\label{fig:ang_dep2} Standard deviations of the difference
    between the computed photo-ionisation rates and an inverse square
    law as a function of ray-to-cell samplings $\Phi_c$ for different
    spatial resolutions.  There is no dependence on the spatial
    resolution, and the accuracy increases as $\sigma \propto
    \Phi_c^{-0.6}$.}
  %\epsscale{1}
\end{figure}

>>>>

Radiation from stars and black holes strongly affects their
surroundings and plays a crucial role in topics such as stellar
atmospheres, the interstellar medium (ISM), star formation, galaxy
formation, supernovae (SNe) and cosmology.  It is a well-studied
problem \citep[e.g.][]{Mathews65, Rybicki, Mihalas84, Yorke86};
however, its treatment in multi-dimensional calculations is difficult
because of the dependence on seven variables -- three spatial, two
angular, frequency, and time.  The non-local nature of the thermal and
hydrodynamical response to radiation sources further adds to the
difficulty.  Depending on the problem of interest some simplifying
assumptions may be made.

@@

\begin{figure}
  \plotone{f1.eps}
  \caption{\label{fig:covering} (a) A two-dimensional illustration of
    the overlap between the beam associated with a ray $\gamma$ and a
    computational cell.  The ray has a segment length of $dr$ passing
    through the cell.  The covering area is denoted by dark grey,
    where the full area $(dr \times L_{\rm pix})$ is colored by the
    dark and light grey.  The photo-ionisation and photo-heating rates
    should be corrected by the ratio of these areas, i.e. the overlap
    fraction $f_c$.  (b) Annotation of quantities used in this
    geometric correction.}
\end{figure}

@@

In an optically thin medium, radiation is only attenuated by geometric
dilution in the local approximation to Equation (\ref{eqn:rteqn}),
i.e. the inverse square law.  With such a simple solution, the tracing
of rays is wasteful, however these rays must be propagated because the
the medium farther away can be optically thick.  Here we describe a
method that minimizes the computational work of ray tracing in the
optically thin regime by exploiting this fact.  Each ray tracks the
total column density $N_{\rm abs}$ and the equivalent total optical
depth $\tau$ traversed by the photon.  If $\tau < \tau_{\rm thin} \sim
0.1$ after the ray exits the cell, we calculate the photo-ionisation
and photo-heating rates directly from the incoming ray instead of the
luminosity of the source.
\begin{equation}
  \label{eqn:optthin}
  k_{\rm ph} = \frac{\sigma_{\rm abs} \; P}{dt_P \; \theta_{\rm pix}}
  \; \frac{r_{\rm cell}}{r_{\rm ray}}.
\end{equation}
Note that the photon number $P$ in the ray has already been
geometrically diluted by ray splitting.  Here $r_{\rm cell}$ and
$r_{\rm ray}$ are the radii from the radiation source to the cell
centre and where the ray exits the cell.  Thus the last factor
corrects the flux to a value appropriate for the cell centre.  The
photo-heating ray is calculated in the same manner as the general
case, $\Gamma_{\rm ph} = k_{\rm ph} (E_{\rm ph} - E_i)$.  This should
only be evaluated once per cell per radiation source.  No photons are
removed from the ray.  With this method, we only require one ray
travel through each cell where the gas is optically thin, thus
reducing the computational expense.

@@

\begin{figure*}
  \epsscale{2}
  \plotone{f2.eps}
  \epsscale{1}
  \caption{\label{fig:MainAlgorithm} Flow chart for the overall
    algorithm of the radiative transfer module in \enzo~that
    illustrates (1) the creation of photon packages, (2) ray tracing,
    (3) the transport of photon packages between AMR grids, and (4)
    coupling with the hydrodynamics.  The ray tracing algorithm, which
    is contained in the ``Trace Rays'' is detailed in Fig.
    \ref{fig:RTalgorithm}.}
\end{figure*}

\begin{figure*}
  \epsscale{2}
  \plotone{f3.eps}
  \epsscale{1}
  \caption{\label{fig:RTalgorithm} Flow chart for the ray tracing
    algorithm for one photon passing through a grid.  Note that only
    one step is needed in the routine to adaptively split rays.  The
    remainder is a typical ray tracing method.}
\end{figure*}

@@

\begin{figure}
  \plottwo{f4a.eps}{f4b.eps}
  \caption{\label{fig:dtsmooth} Radiative transfer adaptive timestep
    in shadowing test (Test 3; \S\ref{sec:test3}) while restricting
    the neutral fraction change to 5\% in the ionisation front.  The
    unmodified timestep (top) is slightly more noisy and the minima
    are more prominent than the timestep computed with a running
    average of the last two timesteps (bottom).  The points show every
    tenth timestep taken into account for the running average.  The
    sawtooth behavior is created by the ionisation front advancing
    into the next neutral cell in the overdensity.}
\end{figure}

@@

\begin{figure}
  \plottwo{f5a.eps}{f5b.eps}
  \caption{\label{fig:test1_ifront} Test 1 (\hii region
    expansion with a monochromatic spectrum of 13.6 eV).  Top:
    Radially averaged profile of neutral (solid) and ionised (dashed)
    fraction at 10, 30, 100, and 500 Myr.  Bottom: Evolution of the
    calculated (top, dashed) and analytical (top, solid) Str\"{o}mgren
    radius.  The ratio of these radii are plotted in the bottom panel.}
\end{figure}

\begin{figure}
  %\epsscale{0.8}
  \plotone{f6.eps}
  \caption{\label{fig:test1_HI} Test 1 (\hii region expansion
    with a monochromatic spectrum of 13.6 eV). Slice of neutral
    fraction at the origin at 500 Myr.}
  %\epsscale{1.0}
\end{figure}

@@

\begin{figure}
  \plotone{f7.eps}
  \caption{\label{fig:test2_1a} Test 2. (\hii region expansion with a
    $T=10^5$ K blackbody spectrum).  Radially averaged profile of
    neutral (solid) and ionised (dashed) fraction at 10, 30, 100, and
    500 Myr.}
\end{figure}

\begin{figure}
  \plotone{f8.eps}
  \caption{\label{fig:test2_1b} Test 2. (\hii region expansion with a
    $T=10^5$ K blackbody spectrum).  Evolution of the average neutral
    fraction.}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f9}
  \epsscale{1}
  \caption{\label{fig:test2_2} Test 2. (\hii region expansion
    with a $T=10^5$ K blackbody spectrum).  Top: Slices through the
    origin of neutral fraction at 10 and 100 Myr.  Bottom: Slices of
    temperature at 10 and 100 Myr.}
\end{figure*}

\begin{figure}
  \plotone{f10}
  \caption{\label{fig:test2_3} Test 2. (\hii region expansion
    with a $T=10^5$ K blackbody spectrum).  Top: Evolution of the
    radius of the simulated ionisation front (dashed) and analytical
    (solid) Str\"{o}mgren radius.  Bottom: The ratio of the calculated
    and analytical Str\"{o}mgren radius.}
\end{figure}

@@

\begin{figure}
  \plottwo{f11a}{f11b}
  \caption{\label{fig:test3_1} Test 3. (I-front trapping in a dense
    clump and shadowing).  Line cut from the point source through the
    middle of the dense clump at $t = 1, 3, 5, 15$ Myr.  of the
    average neutral fraction (top) and temperature (bottom) of the
    clump.}
\end{figure}

\begin{figure}
  %\epsscale{0.85}
  \plotone{f12}
  \caption{\label{fig:test3_2} Test 3. (I-front trapping in a dense
    clump and shadowing).  Evolution of the average ionised fraction
    (top) and temperature (bottom) of the overdense clump.}
  %\epsscale{1}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f13}
  \epsscale{1}
  \caption{\label{fig:test3_3} Test 3. (I-front trapping in a dense
    clump and shadowing).  Clockwise from upper left: Slices through
    the origin of neutral fraction (1 Myr), temperature (1 Myr),
    temperature (15 Myr), and neutral fraction (15 Myr).}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.75}
  \plotone{f14}
  \caption{\label{fig:test4_1} Test 4. (Multiple cosmological
    sources).  Evolution of the mass- (dashed) and volume-averaged
    (solid) ionised fraction.}
  %\epsscale{1}
\end{figure}

\begin{figure*}
  %\epsscale{0.9}
  \epsscale{2}
  \plotone{f15}
  \epsscale{1}
  \caption{\label{fig:test4_2} Test 4. (Multiple cosmological
    sources).  Top: Slices through the origin of neutral fraction at
    50 and 200 kyr at the coordinate $z = z_{\rm box}/2$.  Bottom:
    Slices of temperature at 50 and 200 kyr.  No smoothing has been
    applied to the images.}
  %\epsscale{1}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.8}
  \plotone{f16}
  \caption{\label{fig:test5_1} Test 5. (\hii region in a uniform
    medium).  Top: Growth of the computed ionisation front radius at
    an ionised fraction $x_e = 0.5$ (dashed) and at a temperature $T =
    10^4$ K (dotted) compared to the analytical estimate [solid;
    equation (\ref{eqn:test5_r})].  Bottom: The ratio of the computed
    ionisation front radii to the analytical estimate.} 
  %\epsscale{1}
\end{figure}

\begin{figure}
  \plotone{f17}
  \caption{\label{fig:test5_2} Test 5. (\hii region in a uniform
    medium).  Clockwise from the upper left: Radial profiles of
    density, temperature, ionised fraction, and pressure at times $t
    =$ 10, 200, and 500 Myr.}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f18}
  \epsscale{1}
  \caption{\label{fig:test5_3} Test 5. (\hii region in a uniform
    medium).  Clockwise from the upper left: Slices through the origin
    of ionised fraction, neutral fraction, temperature, and density at
    time $t =$ 500 Myr.}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.8}
  \plotone{f19}
  \caption{\label{fig:test6_1} Test 6. (\hii region in a 1/$r^2$
    density profile).  Top: Growth of the computed ionisation front
    radius as $T=10^4$ K and $x_e = 0.5$ as definitions for the front.
    Bottom: Velocity of the ionisation front, computed from outputs at
    0.5 Myr intervals.  The velocity is calculated from $r_{\rm IF}$,
    whose coarse time resolution causes the noise seen in $v_{\rm
      IF}$.  It is smooth within the calculation itself.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  \plotone{f20}
  \caption{\label{fig:test6_2} Test 6. (\hii region in a 1/$r^2$
    density profile).  Clockwise from the upper left: Radial profiles of
    density, temperature, ionised fraction, and pressure at times $t
    =$ 3, 10, and 25 Myr.}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f21}
  \epsscale{1}
  \caption{\label{fig:test6_3} Test 6. (\hii region in a 1/$r^2$
    density profile).  Clockwise from the upper left: Slices through the origin
    of ionised fraction, neutral fraction, temperature, and density at
    time $t =$ 25 Myr.}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.9}
  \plotone{f22}
  \caption{\label{fig:test7_1} Test 7. (Photo-evaporation of a dense
    clump).  Line cuts from the point source through the middle of the
    dense clump at $t = 1, 10, 50$ Myr of (clockwise from the upper
    left) density, temperature, pressure, and neutral fraction.}
  %\epsscale{1}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f23}
  \epsscale{1}
  \caption{\label{fig:test7_2} Test 7. (Photo-evaporation of a dense
    clump).  Clockwise from the upper left: Slices through the clump
    centre of neutral fraction, pressure, temperature, and density at
    time $t =$ 10 Myr.}
\end{figure*}

\begin{figure*}
  \epsscale{2}
  \plotone{f24}
  \epsscale{1}
  \caption{\label{fig:test7_3} Test 7. (Photo-evaporation of a dense
    clump).  Same as Fig. \ref{fig:test7_1} but at $t = 50$ Myr.}
\end{figure*}

@@

\begin{figure*}
  \epsscale{2}
  \plotone{f25}
  \epsscale{1}
  \caption{\label{fig:test8_1} Test 8. (Champagne flow from a dense
    clump).  Slices of density through the initial clump centre in the
    x-y plane (top) and x-z plane (bottom) at $t = 10, 40, 100, 150$
    kyr.  Notice the instabilities that grow from perturbations
    created while the \hii region is contained in the dense
    clump.}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.75}
  \plotone{f26a}
  \plotone{f26b}
  \plotone{f26c}
  \caption{\label{fig:test9_thick} Test 9. (Irradiated Rayleigh-Taylor
    instability; optically thick case).  Slices at $y=0$ of density
    (top), temperature (middle), and electron fraction (bottom).  The
    source turns on at $t=0$.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  %\epsscale{0.75}
  \plotone{f27a}
  \plotone{f27b}
  \plotone{f27c}
  \caption{\label{fig:test9_thin} Same as Fig. \ref{fig:test9_thick}
    but for the optically thin case.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure*}
  \epsscale{2}
  \plotone{f28}
  \epsscale{1}
  \caption{\label{fig:test10} Test 10 (Photo-evaporation of a
    blastwave).  Slices of density (top) and temperature (bottom) at
    $t = 2.5, 3, 5, 7.5$ Myr in the $x-z$ plane.  As the R-type
    ionisation front propagates through the blastwave centre,
    instabilities grow from the slightly inhomogeneous hot and rarefied
    medium.  Note that the dense shell of the blastwave also creates
    dense inward fingers in the ionisation front shock.}
  %\epsscale{1}
\end{figure*}

@@

\begin{figure*}
  %\epsscale{1}
  \epsscale{2}
  \plotone{f29}
  \epsscale{1}
  \caption{\label{fig:test11} Test 11 (Collimated radiation from a
    dense clump).  Slices of density (top) and temperature (bottom) at
    $t = 0.1, 3.25, 10.75, 23.25$ Myr.  The conical \hii region
    drives shocks transversely into the overdense sphere and creates
    polar champagne flows.  The ambient medium is heated to $T \sim 3
    \times 10^4$ K as the ionisation front passes the
    constant-pressure cloud surface.  The ionisation front changes
    from D-type to R-type after it enters the ambient medium.}
  %\epsscale{1}
\end{figure*}

@@

\begin{figure}
  %\epsscale{1.15}
  \plottwo{f30a}{f30b}
  \caption{\label{fig:test12_kph} Test 12 (Time variations of the
    source luminosity).  Top: Slice of the photo-ionisation rate
    \kph~through the origin.  The source has a duty cycle of 0.5 Myr,
    and the box has a light crossing time of 3.3 Myr.  The shells of
    high \kph~originate from radiation that was emitted when the
    source was at its peak luminosity, illustrating the
    time-dependence of the radiative transfer equation.  Bottom:
    Radial profile \kph~with the inverse square law overplotted.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  %\epsscale{0.8}
  \plotone{f31}
  \caption{\label{fig:test12_profiles} Test 12 (Time variations of the
    source luminosity).  Radial profiles of (clockwise from upper
    left) density, temperature, pressure, and ionised fraction at $t =
    50, 100, 150, 200, 250$ Myr.  In this problem, the time variations
    in the source have little effect on the overall \hii region
    expansion.  Inside the ionisation front at $t = 50$ Myr, there are
    small density perturbations that are created by the variable
    source that are later smoothed out over a sound crossing time.}
  %\epsscale{1}
\end{figure}

\begin{figure*}
  \epsscale{2}
  \plotone{f32}
  \epsscale{1}
  \caption{\label{fig:test14_1} Test 14. (\hii region with MHD).
    Left to right: slices of density at $t = 0.18, 0.53, 1.58$ Myr in
    the x-y plane.  The streamlines show the magnetic field.}
\end{figure*}

\begin{figure*}
  \epsscale{2}
  \plotone{f33}
  \epsscale{1}
  \caption{\label{fig:test14_2} Test 14. (\hii region with
    MHD). Slices of density (top) and the x-component of the magnetic
    field (bottom) in the y-z plane at $t = 0.18, 0.53, 1.58$ Myr
    (left to right).}
\end{figure*}

@@

\begin{figure}
  %\epsscale{0.75}
  \plotone{f34}
  \caption{\label{fig:dx_dep1} Growth of the ionisation front radius,
    compared to the analytical radius, in Test 1 with varying spatial
    resolutions.  At resolutions of $16^3$ and $32^3$, the ionisation
    front is underestimated for the first $\sim25$ Myr but converges
    within 0.5\% of the higher resolution runs.}
  %\epsscale{1}
\end{figure}

@@

\begin{figure}
  %\epsscale{1}
  %\epsscale{2}
  \plotone{f35}
  \epsscale{1}
  \caption{\label{fig:ang_dep1} Variations in the photo-ionisation
    rates for different ray-to-cell samplings $\Phi_c$.  The colormap
    only spans a factor of 3 to enhance the contrast.  In comparison,
    the photo-ionisation rate actually spans 4 orders of magnitude in
    this test.}
  %\epsscale{1}
\end{figure}

\begin{figure}
  %\epsscale{0.75}
  \plotone{f36}
  \caption{\label{fig:ang_dep2} Standard deviations of the difference
    between the computed photo-ionisation rates and an inverse square
    law as a function of ray-to-cell samplings $\Phi_c$ for different
    spatial resolutions.  There is no dependence on the spatial
    resolution, and the accuracy increases as $\sigma \propto
    \Phi_c^{-0.6}$.}
  %\epsscale{1}
\end{figure}
